---
title: "Untitled"
output: html_document
---

```{r setup}
library(tidyverse)
```


# 1. Relative Path

Shorthand	Meaning

  ~	Home directory
  .	Current working directory
  ..	One directory up from current working directory
  ../..	Two directories up from current working directory

ex.
abs path = "/Users/jeffgoldsmith/Dropbox/Work/Teaching/P8105/fall_20209/p8105.github.io/data/FAS_litters.csv"

relative path = "./data/FAS_litter.csv"


# 2. load csv

package::function



```{r}
litters_data = read_csv(file = "./Data/FAS_litters.csv")

#get names
names(litters_data)

#convert names to all small lettercase
litters_data = janitor::clean_names(litters_data)   
# use janitor library, clean_names function, no need to import library
names(litters_data)
```


# 3. look at data

```{r}
# overview
litters_data

#back end, last 5 rows
tail(litters_data, 5)

#dataset summary
skimr::skim(litters_data)

#pop-out winodw view data
view(litters_data)
```



# 4. Read func

col_names: 
usually TRUE. 
If FALSE, column names are X1, X1, … . You can also supply column names.

na: string vector containing character expressions for missing values.

skip: number of rows to skip before reading data.

```{r}
# skip 10 rows, set variable names as X1....Xn
litters_data = read_csv(file = "./Data/FAS_litters.csv", skip = 10, col_names = FALSE)
head(litters_data)
```

# 5. Parse Columns

1. col_types = cols(x1 = _type1, x2 = _type2,....):
specify each cols' variable type


```{r}
litters_data = read_csv(file = "./data/FAS_litters.csv",
  col_types = cols(
    Group = col_character(),
    `Litter Number` = col_character(),
    `GD0 weight` = col_double(),
    `GD18 weight` = col_double(),
    `GD of Birth` = col_integer(),
    `Pups born alive` = col_integer(),
    `Pups dead @ birth` = col_integer(),
    `Pups survive` = col_integer()
  )
)
head(litters_data)
```

* shorthand variable type:
(col_types = "ccddiiiddcc....")

```{r}
litters_data = read_csv(file = "./data/FAS_litters.csv",
                        col_types = "ccddiiii")
litters_data
```


## 1. example 
Repeat the data import process above for the file FAS_pups.csv. Make sure the column names are reasonable, and take some quick looks at the dataset. What happens if your specifications for column parsing aren’t reasonable (e.g. character instead of double, or vice versa)?

```{r}
pups_data = read_csv(file = "./data/FAS_pups.csv", 
                     col_types = "ciiiii")
pups_data
head(pups_data, 20)
tail(pups_data, 20)

skimr::skim(pups_data)

```






# 6. Other format read

read_table:
all other tyeps

read_excel
excel type, with sheet option
                 range option
                 

```{r}
library(readxl)
mlb11_data = read_excel("data/mlb11.xlsx", n_max = 20)
head(mlb11_data, 15)
```


# 7. Base R

read.csv
read.table

* produces tibble, prevent unwarranted change

```{r}
pups_base = read.csv(file = "./data/FAS_pups.csv")
pups_readr = read_csv(file = "./data/FAS_pups.csv")

skimr::skim(pups_base)
skimr::skim(pups_readr)

pups_base
pups_readr

#access Sex variable
pups_base$S
pups_readr$S       # readr version can't read by $S
```



# 8. export data

write_table























